On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-08-15 manuscripts/PNAS/chazthm_pnas.tex

A vector $\mathbf{x}$ is said to be \emph{supported} in $S \subseteq [m]$ when $\mathbf{x} \in \text{\rm span}\{\mathbf{e}_j: j\in S\}$, where $\mathbf{e}_j$ are the standard basis in $\mathbb R^m$. 
%Denote by $\mathbf{x}^J$ the subvector formed from the entries of $\mathbf{x}$ indexed by $J$. 
A set of $k$-sparse vectors is said to be in \emph{general linear position} when any $k$ of them are linearly independent. The following is a precise statement of our main result.  Only the quantity $C_1 = C_1(\mathbf{A}, \{\mathbf{x}_i\}_{i=1}^N, \mathcal{H})$ is undefined (until Sec.~\ref{DUT}, Eq.~\eqref{Cdef1}).
%=======
%where we write $L_{k}$ when $\mathcal{H} = {[m] \choose k}$.\footnote{We note that $1 - \sqrt{k} L_k(\mathbf{M})$ is known as the asymmetric lower restricted isometry constant \cite{Blanchard2011}.} Clearly, for any $\mathbf{M}$ satisfying \eqref{SparkCondition}, we have $L_{k'}(\mathbf{M}) > 0$ for  $k' \leq 2k$. Note that if $\mathcal{H}$ covers $[m]$ (i.e., $\cup \mathcal{H} = [m]$), then $L_2 \geq L_{2\mathcal{H}}$.
% 
%A vector $\mathbf{x}$ is said to be \emph{supported} in $S \subseteq [m]$ when $\mathbf{x} \in \text{\rm span}\{\mathbf{e}_j: j\in S\}$, where $\mathbf{e}_j$ are the standard basis in $\mathbb R^m$. 
%%Denote by $\mathbf{x}^J$ the subvector formed from the entries of $\mathbf{x}$ indexed by $J$. 
%A set of $k$-sparse vectors is said to be in \emph{general linear position} when any $k$ of them are linearly independent. The following is a precise statement of our main result.  Only the quantity $C_1$ is left undefined until later (in Sec.~\ref{DUT}, Eq.~\eqref{Cdef1}).
%>>>>>>> d1407a275dc391dabad7bb62d3e659f9a4ac4624

@@

The total number of $(k-1)$-sparse vectors $\mathbf{\bar x}_i$ can thus not exceed $|\mathcal{H}|(k-1){ \bar m \choose k-1}$. By \eqref{eqn} no more than $|\mathcal{H}|k(k-1){ \bar m \choose k-1}$ vectors $\mathbf{\bar x}_i$ are not $k$-sparse. Since for every $S \in \mathcal{H}$ there are over $(k-1)\left[ {\bar m \choose k} + |\mathcal{H}|k{ \bar m \choose k-1} \right]$ vectors $\mathbf{x}_i$ supported there, it follows that more than $(k-1){\bar m \choose k}$ of them have corresponding $\mathbf{\bar x}_i$ that are $k$-sparse. The result now follows from Thm.~\ref{DeterministicUniquenessCorollary}.
\end{proof}

@@

In the field of theoretical neuroscience in particular, dictionary learning for sparse coding and related methods have recovered characteristic components of natural images \cite{Olshausen96, hurri1996image, bell1997independent, van1998independent} and sounds \cite{bellsejnowski1996, smithlewicki2006, Carlson12} that reproduce response properties of cortical neurons. Our results suggest that this correspondence could be due to the ``universality'' of sparse representations in natural data, an early idea in neural theory \cite{pitts1947}. %Furthermore, they justify the hypothesis of \cite{Isely10} that sparse codes passed through information bottlenecks in the brain are recovered from random projections via (unsupervised) biologically plausible sparse coding (e.g., \cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed, hu2014hebbian}).
Deep learning noise issues: \cite{goodfellow2014explaining}.

>>>>

A vector $\mathbf{x}$ is said to be \emph{supported} in $S \subseteq [m]$ when $\mathbf{x} \in \text{\rm span}\{\mathbf{e}_j: j\in S\}$, where $\mathbf{e}_j$ are the standard basis in $\mathbb R^m$. 
%Denote by $\mathbf{x}^J$ the subvector formed from the entries of $\mathbf{x}$ indexed by $J$. 
A set of $k$-sparse vectors is said to be in \emph{general linear position} when any $k$ of them are linearly independent. The following is a precise statement of our main result.  For sake of brevity, the quantity $C_1 = C_1(\mathbf{A}, \{\mathbf{x}_i\}_{i=1}^N, \mathcal{H})$ is left undefined until Sec.~\ref{DUT} (see \eqref{Cdef1}).
%=======
%where we write $L_{k}$ when $\mathcal{H} = {[m] \choose k}$.\footnote{We note that $1 - \sqrt{k} L_k(\mathbf{M})$ is known as the asymmetric lower restricted isometry constant \cite{Blanchard2011}.} Clearly, for any $\mathbf{M}$ satisfying \eqref{SparkCondition}, we have $L_{k'}(\mathbf{M}) > 0$ for  $k' \leq 2k$. Note that if $\mathcal{H}$ covers $[m]$ (i.e., $\cup \mathcal{H} = [m]$), then $L_2 \geq L_{2\mathcal{H}}$.
% 
%A vector $\mathbf{x}$ is said to be \emph{supported} in $S \subseteq [m]$ when $\mathbf{x} \in \text{\rm span}\{\mathbf{e}_j: j\in S\}$, where $\mathbf{e}_j$ are the standard basis in $\mathbb R^m$. 
%%Denote by $\mathbf{x}^J$ the subvector formed from the entries of $\mathbf{x}$ indexed by $J$. 
%A set of $k$-sparse vectors is said to be in \emph{general linear position} when any $k$ of them are linearly independent. The following is a precise statement of our main result.  Only the quantity $C_1$ is left undefined until later (in Sec.~\ref{DUT}, Eq.~\eqref{Cdef1}).
%>>>>>>> d1407a275dc391dabad7bb62d3e659f9a4ac4624

@@

The total number of $(k-1)$-sparse vectors $\mathbf{\bar x}_i$ can thus not exceed $|\mathcal{H}|(k-1){ \bar m \choose k-1}$. By \eqref{eqn} no more than $|\mathcal{H}|k(k-1){ \bar m \choose k-1}$ vectors $\mathbf{\bar x}_i$ are not $k$-sparse. Since for every $S \in \mathcal{H}$ there are over $(k-1)\left[ {\bar m \choose k} + |\mathcal{H}|k{ \bar m \choose k-1} \right]$ vectors $\mathbf{x}_i$ supported there, it follows that more than $(k-1){\bar m \choose k}$ of them have corresponding $\mathbf{\bar x}_i$ that are $k$-sparse. The result now follows directly from Thm.~\ref{DeterministicUniquenessCorollary}.
\end{proof}

@@

In the field of theoretical neuroscience in particular, dictionary learning for sparse coding and related methods have recovered characteristic components of natural images \cite{Olshausen96, hurri1996image, bell1997independent, van1998independent} and sounds \cite{bellsejnowski1996, smithlewicki2006, Carlson12} that reproduce response properties of cortical neurons. Our results suggest that this correspondence could be due to the ``universality'' of sparse representations in natural data, an early mathematical idea in neural theory \cite{pitts1947}. %Furthermore, they justify the hypothesis of \cite{Isely10} that sparse codes passed through information bottlenecks in the brain are recovered from random projections via (unsupervised) biologically plausible sparse coding (e.g., \cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed, hu2014hebbian}).

Deep learning noise issues: \cite{goodfellow2014explaining}.
